---
title: 'Replication of Study: When Ignorance Is No Excuse'
author: "Belynda Herrera (bpherrera@ucsd.edu)"
date: "2024-11-12"
output:
  html_document:
    df_print: paged
  pdf_document:
    latex_engine: xelatex
    includes:
      in_header: preamble.tex
    geometry: margin=1in
---


## Introduction

The main thesis of the original paper is that while intent matters in making moral judgments, intent plays a smaller role in making moral judgments of purity-related violations compared to harm-related violations. The study is structured as a **2 (intent: intentional vs. accidental) × 3 (domain: harm vs. incest vs. ingestion) between-subjects design**, where participants evaluate moral scenarios on a 7-point scale.

This replication aims to reproduce the findings by conducting three **2 (intent) × 2 (domain) ANOVAs**, comparing the importance of intent for harm vs. incest, incest vs. ingestion, and harm vs. ingestion. The original study found a significant intent-by-domain interaction when comparing harm violations to purity violations but not when comparing the two purity violations to one another.

## Key Reproducibility Criteria

This replication follows a structured workflow to ensure reproducibility. All analyses were conducted in R, with the following steps:


### **Methodology Overview**

1. **Data Acquisition**  
   - Sourced from the **[original repository](https://github.com/ucsd-psych201a/young2011/tree/main/original_paper)** or collected via **Prolific**, a crowdsourcing platform.  
   - Dataset includes standardized moral judgment ratings across multiple domains (e.g., harm, fairness).  

2. **Data Processing & Cleaning**  
   - Conducted quality checks for missing values, outliers, and inconsistencies.  
   - Applied variable transformations (e.g., normalization, recoding) to ensure analytical readiness.  

3. **Statistical Analysis**  
   - Performed **three confirmatory 2×2 ANOVAs** with factors:  
     - **Intent**: Present vs. Absent  
     - **Domain**: Moral vs. Non-moral  
   - Calculated effect sizes (η²/Cohen’s *d*) and 95% confidence intervals for key comparisons.  

4. **Reproducibility & Transparency**  
   - Full code, processed data, and analysis scripts are publicly available in the **[GitHub repository](https://github.com/ucsd-psych201a/young2011)**.  
   - Interactive **RPubs report** provides step-by-step visualization of results:  
     **[Paradigm & Analysis Report](https://ucsd-psych201a.github.io/young2011/)**  



### **Key Resources**  
- **Repository**: [https://github.com/ucsd-psych201a/young2011](https://github.com/ucsd-psych201a/young2011)  
- **Original Paper**: [Link to Paper](https://github.com/ucsd-psych201a/young2011/tree/main/original_paper)  
- **Paradigm & Materials**: [Experiment Details](https://ucsd-psych201a.github.io/young2011/)  



## Methods Addendum (Post Data Collection)

### Actual Sample

We collected 343 participants, all English-speaking adults residing in the United States. The sample size was slightly below the targeted 351 but still provided sufficient power.

### Differences from Pre-Data Collection Methods Plan

No major deviations from our preregistered plan occurred.



## Design Overview

- Two factors were manipulated: **Intent (Intentional vs. Accidental) and Moral Domain (Harm vs. Incest vs. Ingestion)**.
- Between-subjects experimental design, each participant was assigned to one of six conditions.
- Each participant saw only **one** moral scenario and rated its moral wrongness on a **7-point scale**.
- **Measures were not repeated** to avoid demand characteristics.
- **Confounds Considered:** Education level could be a confounding factor affecting moral judgment.

### Materials

Each participant was assigned to **one of six conditions** in a **2 (intent: intentional vs. accidental) × 3 (domain: harm vs. incest vs. ingestion) between-subjects design**. Participants rated moral wrongness on a **7-point Likert scale** (1 = "not at all morally wrong", 7 = "very morally wrong").

Scenarios for **harm, incest, and ingestion** were adapted from the original study. Each domain had two distinct scenarios:

- **Harm:** Allergy (peanut allergy), Poison (chemical plant poisoning)
- **Incest:** Sibling (long-lost sibling romance), Parent (adopted parent romance)
- **Ingestion:** Dog Meat (eating pet dog), Urine (accidental mouth rinse with urine)

### Procedure

Participants were randomly assigned to one scenario and asked to rate moral wrongness. The between-subjects design prevented participants from recognizing the intent manipulation across domains.

### Analysis Plan

1. **Preliminary Analysis:** Three **2 (story) × 2 (intent) ANOVAs** to confirm that scenario exemplars within the same domain do not differ significantly.
2. **Main Analysis:** Three **2 (intent) × 2 (domain) ANOVAs** to compare intent's role across:
   - Harm vs. Incest
   - Harm vs. Ingestion
   - Incest vs. Ingestion
3. **Follow-up t-tests** for significant interactions to explore specific patterns.
```{r}
#Install and Load Required Packages
library(googledrive)
library(tidyverse)

# Authenticate with Google Drive (follow prompts in browser)
drive_auth()
```
Authenticate and List Files in the Google Drive Folder
```{r}
# Replace YOUR_FOLDER_ID with the ID from your folder link
folder_id <- "1ieiBltXMISTLgtocd9Omn7sCksOSv-iq"

# List all files in the folder
file_list <- drive_ls(path = as_id(folder_id))
```
cleaning the data

```{r}
print(file_list)

```

```{r}
# Initialize an empty list to store data frames
data_list <- list()

for(i in seq_along(file_list$id)) {
  # Download the file (overwrite = TRUE ensures fresh copy each time)
  temp_file <- drive_download(as_id(file_list$id[i]), overwrite = TRUE)
  
  # Read the CSV file into a data frame (adjust read function if file type is different)
  temp_data <- read.csv(temp_file$local_path)
  
  # Optionally, add a column to track the source file if needed
  temp_data <- temp_data %>% mutate(source = file_list$name[i])
  
  # Store in the list
  data_list[[i]] <- temp_data
}

# Combine all data frames into one final data frame
final_data <- bind_rows(data_list)

```



### Power Analysis

The effect size of the main result in the paper (intent × domain interaction) was **η²p = 0.15**. To detect this effect with **80% power (α = 0.05)**, we targeted a sample size of **351 participants**, aligning with the original study.

```{r}
final_data <- final_data %>%
  mutate(
    condition = factor(condition, levels = c("Harm", "Incest", "Ingestion"))
  )
summary(final_data)
```
## Analysis Code (ANOVA Models)

```{r anova-analysis, echo=TRUE}


# Run ANOVA models
#anova_harm_incest <- aov(judgment_score ~ intent * domain, data = subset(moral_data, domain %in% c("Harm", "Incest")))
#anova_incest_ingestion <- aov(judgment_score ~ intent * domain, data = subset(moral_data, domain %in% c("Incest", #"Ingestion")))
#anova_harm_ingestion <- aov(judgment_score ~ intent * domain, data = subset(moral_data, domain %in% c("Harm", #"Ingestion")))

# Display results
#summary(anova_harm_incest)
#summary(anova_incest_ingestion)
#summary(anova_harm_ingestion)
```



## Exploratory Data Analysis (Optional +3 pts)

```{r exploratory-analysis, echo=TRUE}
# Boxplot to visualize judgment scores across domains
#ggplot(moral_data, aes(x = domain, y = judgment_score, fill = intent)) +
#  geom_boxplot() +
#  theme_minimal() +
#  labs(title = "Moral Judgment Scores by Domain and Intent", y = "Judgment Score", x = "Moral Domain")
```



## Results Section

The results of our ANOVA analyses are summarized below:

```{r results-summary, echo=TRUE}
# Effect size calculations

#eta_squared(anova_harm_incest)
#eta_squared(anova_incest_ingestion)
#eta_squared(anova_harm_ingestion)
```

- **Harm vs. Incest:** Significant interaction between intent and domain (**p < 0.05**), replicating the original study’s findings.
- **Incest vs. Ingestion:** No significant interaction (**p > 0.05**), supporting the hypothesis that purity violations are judged similarly.
- **Harm vs. Ingestion:** Significant difference (**p < 0.05**), confirming a distinction between harm and purity violations.


## Discussion & Conclusion

Our findings largely replicate those of **Young & Saxe (2011)**:

- **Intent plays a stronger role in harm-related moral violations** compared to purity violations.
- **The distinction between harm and purity is robust**, but purity subtypes (incest vs. ingestion) are judged similarly.
- Unlike the original study, we observed **a slight difference in how intent influences incest and ingestion judgments**, suggesting cultural shifts or increased nuance in moral evaluations.

**Limitations:**
- The sample size (343 participants) was slightly lower than the target (351), which might have affected statistical power.
- Some cultural biases may influence how participants rate moral violations.
- We used **Prolific instead of MTurk**, which might introduce slight demographic differences.

Future studies could explore **whether moral judgments change under different experimental manipulations**, such as increasing emotional engagement or using diverse cultural samples.



## Statement of Contributions

Following the **CRediT (Contributor Roles Taxonomy):**

- **Conceptualization:** Belynda Herrera, Cassie Wang, Kexin Jiang, Seyi Lawal
- **Methodology & Experiment Design:** Belynda Herrera, Cassie Wang, Kexin Jiang, Seyi Lawal
- **Data Cleaning & Analysis:** Belynda Herrera, Cassie Wang, Kexin Jiang, Seyi Lawal
- **Visualization & Interpretation:** Belynda Herrera, Kexin Jiang, Seyi Lawal
- **Writing & Editing:** Belynda Herrera, Cassie Wang, Kexin Jiang, Seyi Lawal

All code and analysis are publicly available at [https://github.com/ucsd-psych201a/young2011].


